---
title: pcIRT
author: 'Christine Hohensinn'
date: '2018-03-02'
slug: pcirt

weight:  10

summary: "pcIRT estimates the Multidimensional Polytomous Rasch model and the Continuous Rating Scale model"
image_preview: 'headers/iccMPRM.png'

categories:
  - R
tags: ["R", "pcIRT", "IRT", "Rasch model"]
---

The R package *pcIRT* provides functions that estimate the following IRT models:

- **Multidimensional Polytomous Rasch model (MPRM)** (Rasch, 1961)

MPRM is a multidimensional IRT model for multicategorical data. For each response category, a latent dimension is assumed. The item characteristic curves of an example looks like this:


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(pcIRT)

```


```{r create_data}
itpar <- matrix(rnorm(3*5, 0,1), 3, 5)
itpar2 <- cbind(itpar,rowSums(itpar)*(-1))

simdat <- simMPRM(itpar2, persons = 500, seed=3324)
mprm_erg <- MPRM(simdat$datmat)

pp1 <- seq(-5,5, by=0.1)
pp2 <- seq(-5,5, by=0.1)


object <- mprm_erg
items <- 6
l <- 1
p1 <- 1

#aus pcIRT package
    pp_comb <- expand.grid(theta1 = pp1, theta2 = pp2)
    m1 <- exp(pp_comb[,1] - object$itempar[1,l])
    m2 <- exp(pp_comb[,2] - object$itempar[2,l])
    
    n <- rowSums(cbind(m1,m2, exp(0)))
    respc1 <- m1/n
    respc2 <- m2/n
    r2 <- cbind(respc1, respc2)
    respc3 <- 1-rowSums(r2)
    
    mat <- cbind(respc1, respc2, respc3)
    colnames(mat) <- paste0("cat", 1:3)

      zmat <- matrix(mat[,p1], ncol=length(pp1))
      row.names(zmat) <- pp1
      colnames(zmat) <- pp2

      
      zmat2 <- matrix(mat[,2], ncol=length(pp1))
      zmat3 <- matrix(mat[,3], ncol=length(pp1))


      

```

```{r plot}

    axx <- list(
      title = "theta cat 1",
      backgroundcolor="#e6e6e6",
      gridcolor="#cccccc",
      showbackground=TRUE,
      zerolinecolor="#4d4d4d"
    )
    axy <- list(
      title = "theta cat 2",
      backgroundcolor="#e6e6e6",
      gridcolor="#cccccc",
      showbackground=TRUE,
      zerolinecolor="#4d4d4d"
    )
    axz <- list(
      title = "resp. prob.",
      backgroundcolor="#e6e6e6",
      gridcolor="#cccccc",
      showbackground=TRUE,
      zerolinecolor="#4d4d4d"
    )
    
    
    plot_ly(showscale = FALSE, colors = c('#ff9900','#ff0000' )) %>%
      add_surface(x = pp1, y = pp1, z = zmat, opacity = 0.8) %>%
      add_surface(x = pp1, y = pp1, z = zmat2, opacity = 0.7) %>%
      add_surface(x = pp1, y = pp1, z = zmat3, opacity = 0.7) %>%
      config(displayModeBar = F) %>%
      layout(scene = list(xaxis = axx, yaxis = axy, zaxis = axz))



```



- **Continuous Rating Scale model (CRSM)** (Müller, 1987)

CRSM is an extension of the Rating Scale model. It can be used to analyze items with a visual analogue scale for example.

- **Dichotomous Logistic Rasch model** (Rasch, 1960) 


By defining the design matrix, restricted Rasch models (e.g. the Linear Logistic Rasch model) can be estimated, too. 


### Download
pcIRT is available on [CRAN](https://cran.r-project.org/web/packages/pcIRT/index.html). The developer version can be downloaded from [my Github Site](https://github.com/christinehohensinn/pcIRT).

### Publication
pcIRT (especially the application and usage of the MPRM) is described and illustrated in [this paper](https://www.jstatsoft.org/article/view/v084c02)


